---
title: "RNA seq Report"
author: "Niel Infante"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=F)


# Error logging
log <- file(snakemake@log[[1]], open="wt")
sink(log)
sink(log, type="message")

# Load needed packages
library(tidyverse)
library(DESeq2)
library("RColorBrewer")
library(ggrepel)
library(gplots)
library(pheatmap)
library(knitr)

# Read dds
exp <- snakemake@params$exp
dds <- readRDS(paste0("deseq/", exp, "/dds.rds"))

#vsd <- vst(dds, blind=F)

# For testing only
vsd <- varianceStabilizingTransformation(dds, blind=F)


# Read Metadata and source config
meta_file_name <- snakemake@config$metadata_file
meta <- read_tsv(meta_file_name)

snakemake@source(paste0('../deseq/', exp, '/config.R'))

# Read results
res <- read_tsv(paste0('deseq/', exp, '/results.txt'))

# Read CV
cv <- read_tsv(paste0('deseq/', exp, '/basemean_cutoff.txt'))
cut_value <- cv$CV[1]
```

## Results

These are the top 10 most significant genes
```{r results}

kable(res[1:10,])

```

The full list is available in the file results.txt

## Fold Change

```{r sanity}

title=paste(res[1,]$GeneName, "\nFold Change:",res[1,]$log2FoldChange)
plotCounts(dds, gene=res[1,]$GeneID, intgroup = PCA_Group, main=res[1,]$GeneName, 
					 sub=paste('FC:', format(res[1,]$log2FoldChange, digits=2)), pch=19)



title=paste(res[1,]$GeneName, "\nFold Change:",res[1,]$log2FoldChange)
plotCounts(dds, gene=res[1,]$GeneID, intgroup = PCA_Group, main=res[1,]$GeneName, 
					 sub=paste('FC:', format(res[1,]$log2FoldChange, digits=2)), pch=19)



```

Fold change is reported as log2 (Fold Change). This graph give you an idea of what a positive versus negative fold change means.

## Expression Cutoff

```{r expr_cut}
ggplot(res, aes(x=log2(baseMean))) + geom_density() + 
	geom_vline(xintercept=cut_value, color='red') +
	geom_text(aes(x=cut_value, y=.6, label='   Expression Cutoff', hjust='left'), color='red')


```

Genes are filtered if the average expression across all samples is less than the cut off value, in this case `r 2^cut_value`.


## MA Plot

```{r maplot}
plotMA(dds)
```

## Sample Distance

```{r sample_heat}

select <- order(rowMeans(counts(dds,normalized=TRUE)),decreasing=TRUE)[1:100]
hmcol <- colorRampPalette(brewer.pal(9, "GnBu"))(100)

distsRL <- dist(t(assay(vsd)))
mat <- as.matrix(distsRL)
rownames(mat) <- colnames(mat) <- with(colData(dds), paste(Graph_Display))

heatmap.2(mat, trace="none", col = rev(hmcol), margin=c(8, 8), density.info = 'none')

```

This is a heatmap of sample distances. Dark indicates the samples are closer together, based on the expression of the 100 most expressed genes.


## Cluster

```{r cluster}
plot(hclust(dist(t(assay(vsd)))), label=with(colData(dds), paste(Graph_Display)), main='Dendrogram', xlab='', sub='')

```


## PCA
```{r pca}


print(plotPCA(vsd, intgroup=c(PCA_Group)))
```

A Principle Conmponent plot of the data. Each point is a sample.


## PCA with Names
```{r pca_names}

p <- plotPCA(vsd, intgroup=c(PCA_Group))
p <- p + geom_text_repel(aes(x=PC1, y=PC2, 
				label=with(colData(dds), paste(Graph_Display))), 
				point.padding = unit(2,"points"))
print(p)

```

The same PCA plot, with added names



##  Volcano

```{r volcano}

par(pch = 16)
with(res, plot(log2FoldChange, -log10(pvalue), main = "Volcano plot"))
with(subset(res, padj < 0.05), points(log2FoldChange, -log10(pvalue), col = "red"))
with(subset(res, abs(log2FoldChange) > 2), points(log2FoldChange, -log10(pvalue),  col = "orange"))

with(subset(res, padj < 0.05 & abs(log2FoldChange) > 2), points(log2FoldChange,  -log10(pvalue), col = "green"))

# Add legend
legend("topleft", legend = c("FDR<0.05", "|LFC|>2", "both"), pch = 16, col = c("red", "orange", "green"))

```


## Expression Heatmap

```{r pheatmap, message=F, warning=F}

rld <- rlogTransformation(dds)


# Get 15 up, 15 down 

up <- head(order(res$log2FoldChange),15)
up <- res[up,'GeneID']


dn <- head(order(res$log2FoldChange, decreasing = T),15)
dn <- res[dn,'GeneID']

#get <- c(up,dn)
get <- c(unlist(up),unlist(dn))


rld15 <- rld[get]
#rld15 <- rld[rld@rowRanges@partitioning@NAMES %in% get, ]

rn <- data.frame(GeneID=rownames(rld15))
rn <- left_join(rn, res)

rownames(rld15) <- rn$GeneName

mat <- assay(rld15)
mat <- mat - rowMeans(mat)     # Get difference from mean for each gene
df <- as.data.frame(colData(rld15)[,PCA_Group])
names(df) <- PCA_Group

# Need the same names in both the df and the matrix
rownames(df) <- colData(rld15)[,c('Graph_Display')]
colnames(mat) <- colData(rld15)[,c('Graph_Display')]


pheatmap(mat, cutree_rows = 2, cutree_cols = 2, annotation_col = df)

```

Heatmap shows the 15 most upregulated, and 15 most downregulated genes. Color indicates how much the expression of the gene inthe samples is different from the mean expression of that gene.







