## Gene Ontology

Gene ontology analysis was done on genes found to be significantly differentially expressed. The GO analysis was done separately for significantly up regulated genes, significantly down regulated genes, and all significant genes. The background in all cases is genes with enough expression to make it past the initial filters. The analysis was done for Molecular Function, Biological Process, and Cellular Component. Tables and figures were created for all of these analyses, and are available in the GO sub folder. Below are some selected outputs.


```{r go_steup, include=F}

# See if there are any results for molecular function.
# Try up first, then down, then all
# set the appropriate table and picture
showMF <- TRUE

mf_res <- read_tsv(paste0(snakemake@scriptdir, "/../results/", exp, "/GO/all_up_genes_MF_results.txt"))
mf_pic <- paste0(snakemake@scriptdir, "/../results/", exp, "/GO/all_up_genes_MF_bar.png")
mf_message <- "Significantly up regulated genes"

if(dim(mf_res)[1] == 0){
	mf_ref <- read_tsv(paste0(snakemake@scriptdir, "/../results/", exp, "/GO/all_down_genes_MF_results.txt"))
	mf_pic <- paste0(snakemake@scriptdir, "/../results/", exp, "/GO/all_down_genes_MF_bar.png")
	mf_message <- "Significantly down regulated genes"

	if(dim(mf_res)[1] == 0){
		mf_res <- read_tsv(paste0(snakemake@scriptdir, "/../results/", exp, "/GO/all_genes_MF_results.txt"))
		mf_pic <- paste0(snakemake@scriptdir, "/../results/", exp, "/GO/all_genes_MF_bar.png")
		mf_message <- "All significant genes"

		if(dim(mf_res)[1] == 0){
			showMF <- FALSE
			mf_message <- "No significant ontological terms were found for Molecular Function"
		}
	} 
}


# Same for BP
showBP <- TRUE

bp_res <- read_tsv(paste0(snakemake@scriptdir, "/../results/", exp, "/GO/all_up_genes_BP_results.txt"))
bp_pic <- paste0(snakemake@scriptdir, "/../results/", exp, "/GO/all_up_genes_BP_dot.png")
bp_message <- "Significantly up regulated genes"

if(dim(bp_res)[1] == 0){
	bp_ref <- read_tsv(paste0(snakemake@scriptdir, "/../results/", exp, "/GO/all_down_genes_BP_results.txt"))
	bp_pic <- paste0(snakemake@scriptdir, "/../results/", exp, "/GO/all_down_genes_BP_dot.png")
	bp_message <- "Significantly down regulated genes"

	if(dim(bp_res)[1] == 0){
		bp_res <- read_tsv(paste0(snakemake@scriptdir, "/../results/", exp, "/GO/all_genes_BP_results.txt"))
		bp_pic <- paste0(snakemake@scriptdir, "/../results/", exp, "/GO/all_genes_BP_dot.png")
		bp_message <- "All significant genes"

		if(dim(bp_res)[1] == 0){
			showBP <- FALSE
			bp_message <- "No significant ontological terms were found for Biological Process"
		}
	}
}

# Same for CC
showCC <- TRUE

cc_res <- read_tsv(paste0(snakemake@scriptdir, "/../results/", exp, "/GO/all_up_genes_CC_results.txt"))
cc_pic <- paste0(snakemake@scriptdir, "/../results/", exp, "/GO/all_up_genes_CC_GO_graph.png")
cc_message <- "Significantly up regulated genes"

if(dim(cc_res)[1] == 0){
	cc_ref <- read_tsv(paste0(snakemake@scriptdir, "/../results/", exp, "/GO/all_down_genes_CC_results.txt"))
	cc_pic <- paste0(snakemake@scriptdir, "/../results/", exp, "/GO/all_down_genes_CC_GO_graph.png")
	cc_message <- "Significantly down regulated genes"

	if(dim(cc_res)[1] == 0){
		cc_res <- read_tsv(paste0(snakemake@scriptdir, "/../results/", exp, "/GO/all_genes_CC_results.txt"))
		cc_pic <- paste0(snakemake@scriptdir, "/../results/", exp, "/GO/all_genes_CC_GO_graph.png")
		cc_message <- "All significant genes"

		if(dim(cc_res)[1] == 0){
			showCC <- FALSE
			cc_message <- "No significant ontological terms were found for Cellular Components"
		}
	}
}




```
## GO Analysis Molecular Function
`r mf_message`

```{r go_mf_table, message=F, warning=F}
res <- read_tsv(paste0(snakemake@scriptdir, "/../results/", exp, "/GO/all_up_genes_MF_results.txt"))

if(showMF){
	mf_res <- mf_res %>% select(ID, Description, GeneRatio, BgRatio, p.adjust)
	datatable(mf_res, filter='bottom', options = list(pageLength=10, lengthMenu = c(5, 10, 25), scrollX=T)) %>% 
		formatRound('p.adjust', 3)
} 
```

## GO Analysis Molecular Function
`r mf_message`

```{r go_mf_plot, out.width="90%", message=F, warning=F}

if(showMF){
	include_graphics(mf_pic)
}
```


## GO Analysis Biological Processes
`r bp_message`

```{r go_bp_table, message=F, warning=F}

if(showBP){
	bp_res <- bp_res %>% select(ID, Description, GeneRatio, BgRatio, p.adjust)
	datatable(bp_res, filter='bottom', options = list(pageLength=10, lengthMenu = c(5, 10, 25), scrollX=T)) %>% 
		formatRound('p.adjust', 3)
	#	formatRound('baseMean', 0)
	
}

```

## GO Analysis Biological Processes
`r bp_message`

```{r go_bp_plot, out.width="90%", message=F, warning=F}

if(showBP){
	include_graphics(bp_pic)
}


```


## GO Analysis Cellular Components
`r cc_message`

```{r go_cc_table, message=F, warning=F}

if(showCC){
	cc_res <- cc_res %>% select(ID, Description, GeneRatio, BgRatio, p.adjust)
	datatable(cc_res, filter='bottom', options = list(pageLength=10, lengthMenu = c(5, 10, 25), scrollX=T)) %>% 
		formatRound('p.adjust', 3)
	#	formatRound('baseMean', 0)
	
}

```

## GO Analysis Cellular Components
`r cc_message`

```{r go_cc_plot, out.width="95%", message=F, warning=F}

if(showCC){
	include_graphics(cc_pic)
}

```

```{r go_cleanup, include=F}

# Remove variables used, to free up a little space

rm(mf_res)
rm(mf_pic)
rm(bp_res)
rm(bp_pic)
rm(cc_res)
rm(cc_pic)


```