---
title: "Test.rmd"
author: "Niel Infante"
date: "12/2/2019"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

```{r, result='asis'}
#cat(knitr::knit_child('Test_child.rmd', quiet=T))

#cat(knitr::asis_output(knitr::knit_child("Test_child.rmd", quiet=TRUE)))

#cat(knitr::knit_child("child.Rmd", quiet=TRUE, envir=environment()))

```

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)

doChild <- T
childVar <- 'Hey, its me'

```

```{r testing, child=if(doChild) 'Test_child.rmd' else 'empty.rmd'}

```

## Slide with Plot

```{r pressure}
plot(pressure)
```

